---
title: "Esophagus REDCap Treatment Data vs OncoRegimenFinder"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{esophagus-redcap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(rubix)
library(OncoRegimenFinder)

```

```{r, echo=FALSE,eval=TRUE,message=FALSE,results="hide"}
library(fantasia)
conn <- fantasia::connectOMOP()
```

### Esophagus Cohort OncoRegimenFinder Results  
1. Get the Current Esophagus Cohort  
1. Filter oncoregimen_regimen_ingredients Table for Esophagus Cohort

```{r, eval = TRUE, echo=FALSE}

sql <-
pg13::buildJoinQuery(fields = c("patelm9.esophagus_cohort.person_id AS esophagus_person_id",
                                "patelm9.esophagus_cohort.ptid",
                                "patelm9.oncoregimenfinder_regimen_ingredients.*"),
                     schema = "patelm9",
                     tableName = "esophagus_cohort",
                     column = "person_id",
                     joinType = "LEFT",
                     joinOnSchema = "patelm9",
                     joinOnTableName = "oncoregimenfinder_regimen_ingredients",
                     joinOnColumn = "person_id")

cat(sql)

```


```{r, echo=FALSE, eval=TRUE,results='hide'}

esophagus_oncoregimenfinder_results <-
pg13::query(conn = conn,
             sql_statement = sql)

```


```{r, echo = FALSE, eval = TRUE}
esophagus_oncoregimenfinder_results <-
  esophagus_oncoregimenfinder_results %>%
  tibble::as_tibble() %>% 
  dplyr::mutate_all(as.character) %>%
  rubix::normalize_all_to_na() 
print(esophagus_oncoregimenfinder_results)
```

## OncoRegimenFinder Results Summary  
### Missingness  

```{r}
esophagus_oncoregimenfinder_results %>%
  rubix::summarize_variables() %>%
  dplyr::select(-contains("DISTINCT_VALUES"))
```

## REDCap Data  

```{r}
# redcap_data <- fantasia::read_redcap_data("~/GitHub/MSK-patelm9/escritoire/Esophagus/data_with_phi/STGIEsophagogastricS_DATA_2020-07-28_1354.csv")

input <-
fantasia::processRedcapData(data_csv_file = "~/GitHub/MSK-patelm9/escritoire/Esophagus/data_with_phi/STGIEsophagogastricS_DATA_2020-07-28_1354.csv",
                  parsed_metadata_csv_file = "~/GitHub/MSK-patelm9/escritoire/Esophagus/data_without_phi/STGIEsophagogastricSandbox_DataDictionary_2020-06-24_Parsed.csv",
                    identifier_variables = "ptid")

```

The following are Treatment Variables with Text or Selection Values: 
```{r}

identifier_vars <- "ptid"
treatment_vars <-
  c('neoadj_regimen',
  'neoadj_tx_additional_drugs',
  'neoadj_tx_addtl_drugs_2',
  'crt_name',
  'neoadjcrt_tx_additional_drugs',
  'neoadjcrt_tx_addtl_drugs_2',
  'adj_regimen',
  'adj_tx_additional_drugs',
  'adj_tx_addtl_drugs_2',
  'adj_rt_concurrent',
  'adjcrt_tx_additional_drugs',
  'adjcrt_tx_addtl_drugs_2',
  'induction_regimen',
  'periop_neo_regimen',
  'periop_adj_regimen',
  'tx_name',
  'advanced_tx_additional_drugs',
  'advanced_tx_addtl_drugs_2',
  'tx_maintenance_drugs',
  'advanced_tx_addtl_drugsm1_1',
  'advanced_tx_addtl_drugsm1_2',
  'tx_maintenance2_drugs',
  'advanced_tx_addtl_drugsm2_1',
  'advanced_tx_addtl_drugsm2_2',
  'palliative_rt_concurrent',
  'neoadj_tx_base_regimen',
  'neoadjcrt_tx_base_regimen',
  'adj_tx_base_regimen',
  'adjcrt_tx_base_regimen',
  'advanced_tx_base_regimen',
  'regimen_name_m1',
  'regimen_name_m2',
  'hipec_drugs')



```



```{r}
qa <- treatment_vars[!(treatment_vars %in% colnames(input$ProcessedData))]
treatment_vars <- treatment_vars[(treatment_vars %in% colnames(input$ProcessedData))]
```


```{r}

output <-
treatment_vars %>%
  rubix::map_names_set(function(x) input$ProcessedData %>%
                         tibble::as_tibble() %>%
                              rubix::normalize_all_to_na() %>%
                            dplyr::filter_at(vars(all_of(x)),
                                             any_vars(!is.na(.))) %>%
                            dplyr::select(all_of(identifier_vars),
                                          all_of(x),
                                          contains("start"),
                                          contains("stop")) %>%
                         rubix::deselect_if_all_na()) %>%
  purrr::keep(~nrow(.) > 0) %>% 
  # If the df has less than 2 columns, that means there wasn't a date associated with the variable
  purrr::keep(~ncol(.) > 2)

```



```{r}
output2 <-
  output %>%
  purrr::map2(names(output), function(x, y) x %>%
                          tidyr::pivot_longer(cols = c(contains("start"), contains("stop")),
                                              names_to = c("Context",
                                                           "DateType"),
                                              names_pattern = "(^.*)_([start|stop].*$)",
                                              values_to = "Date")) %>%
  purrr::map(function(x) x %>%
                          dplyr::rename_at(vars(2),
                                           ~paste("PV"))) %>%
  dplyr::bind_rows(.id = "VARIABLE")
```

```{r}
output3 <-
  output2 %>% 
  dplyr::mutate(DateType = forcats::as_factor(DateType)) %>% 
  dplyr::mutate(DateType = forcats::fct_collapse(DateType,
                                                 EVENT_START_DATE = c("start", "start_date"),
                                                 EVENT_END_DATE = c("stop", "stop_date"))) %>%
  tidyr::pivot_wider(id_cols = c(!DateType, !Date),
                     names_from = DateType,
                     values_from = Date,
                     values_fn = list(Date = function(x) paste(unique(x), collapse = "\n")))

```

```{r}

esophagus_cohort <- 
  pg13::readTable(conn = conn,
                  schema = "patelm9",
                  tableName = "esophagus_cohort")

output4 <-
  output3 %>%
  dplyr::full_join(esophagus_cohort) %>%
  dplyr::select(all_of(colnames(output3)),
                person_id) %>%
  dplyr::mutate(NORMALIZE_START_DATE = EVENT_START_DATE,
                NORMALIZE_END_DATE = EVENT_END_DATE)


```

```{r}
output5 <- split(output4, output4$person_id)
```

Split by person_id because the esophagus_person_id is the most comprehensive source of truth in terms of the patient population. Person_id is only for regimen matches.


```{r}

esophagus_results2 <-
esophagus_oncoregimenfinder_results %>%
  dplyr::mutate(NORMALIZE_START_DATE = ingredient_start_date,
                NORMALIZE_END_DATE = ingredient_end_date) %>%
  dplyr::mutate_at(vars(person_id), as.integer) 
esophagus_results3 <- split(esophagus_results2, esophagus_results2$esophagus_person_id)

```


```{r}
list(REDCAP = output5,
     OMOP = esophagus_results3) %>%
  purrr::transpose() %>% 
  purrr::map(dplyr::bind_rows, .id = "DATA_SOURCE")
```




```{r,echo=FALSE,eval=TRUE,message=FALSE,results="hide"}
fantasia::dcOMOP(conn = conn,
                 remove = TRUE)
```


