---
title: "Introduction to OncoRegimenFinder"  
author: "Meera Y. Patel, MD"
description: > 
    Learn how to run OncoRegimenFinder 
output: 
    rmarkdown::html_vignette:
      toc: true
      df_print: kable
vignette: >
  %\VignetteIndexEntry{Introduction to OncoRegimenFinder}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  tidy = 'styler',
  collapse = TRUE,
  comment = "#>"
)
```

# Overview  
The goal of OncoRegimenFinder is to write tables that derive antineoplastic drug combinations from every patient in the OMOP CDM instance. The execution occurs in 2 overarching steps:  
1. Subsetting the Drug Exposure table for antineoplastic drugs   
2. Executing an algorithm that derives antineoplastic drug combinations based on overlaps in the window of time created by adding and subtracting the `date_lag_input` in days from the start date of the drug exposure. This initial resultset is then joined onto itself for a number of iterations determined by `regimen_repeats`, the maximum number of drugs in derived drug combinations.  

It is important to note that the procedures shown below will run an algorithm with a set `date_lag_input` and `regimen_repeats` across the entire OMOP CDM Drug Exposure Table. Though the ultimate goal is to enable retrieval of derived drug combinations from the final `writeDatabaseSchema.regimenIngredientTable` via a join on the `person_id` field, specific cohorts may require different `date_lag_input` and `regimen_repeats`. For example, esophageal carcinoma patients tend to have regimens executed in blocks of 2 weeks instead of 30 days and the algorithm may return more accurate results with a `date_lag_input` of 15 days. In special cases such as this it, a second run of OncoRegimenFinder with modified parameters should be run with customized table names.  

# General Procedure Notes  
* If any of the tables written in this package already exist in the given schema, the existing table is renamed with the system date appended in the format {table}_YYYY_mm_dd. If {table}_YYYY_mm_dd already exists, the process is skipped and the main table (ie {table}) is dropped.  

# Setup  
Setup involves writing an Ingredient Exposure Table and a Vocabulary Table that serve as the data sources for the remainder of the OncoRegimenFinder procedure.  

## Setup Requirements  
1. A database connection object `conn` to your OMOP CDM instance   
1. `writeDatabaseSchema` destination schema with write privileges  

## Outputs   
The objective of the setup is to map all data sources to RxNorm Ingredients, which will serve as the point of alignment between Drug Exposures and a Regimen. This involves creating the following tables in `writeDatabaseSchema`:  
1. Ingredient Exposure Table (`drugExposureIngredientTable`): Drug Exposure Table with an additional `ingredient_concept_id` field that references the RxNorm Ingredient that corresponds to the `drug_concept_id` in the OMOP Vocabulary.   
1. Vocabulary Table (`vocabularyTable`): In a similar fashion as the Ingredient Exposure Table, HemOnc Regimens are mapped to mapped to 1 or more RxNorm Ingredients.  

## Maintenance   
* The Ingredient Exposure Table should be rewritten according to the cadence in which the Drug Exposures Table source is updated.  
* The Vocabulary Table should be rewritten whenever there is an update to HemOnc, RxNorm, or RxNorm Extension in the OMOP Vocabulary.  

## Procedure   
### Load Packages  

```{r pkgsetup, echo=TRUE,eval=TRUE,message=FALSE,results="hide", warning=FALSE}
library(OncoRegimenFinder)
library(tidyverse)
```


```{r secretpkgsetup, echo=FALSE,eval=TRUE,message=FALSE,results="hide",warning=FALSE}
library(fantasia)
conn <- fantasia::connectOMOP()
```

### Set Parameters   
* `cdmDatabaseSchema`: schema housing the Concept and Drug Exposure Table  
* `writeDatabaseSchema`: schema to write the Ingredient Exposure and Vocabulary Table to  
* `drugExposureIngredientTable`: name of the Ingredient Exposure Table  
* `vocabularyTable`: name of the Vocabulary Table  
        
```{r setupparameters, echo=TRUE,eval=TRUE}
# Schema for the Concept and Drug Exposure Tables
cdmDatabaseSchema <- "omop_cdm_2"

# Output Schema and Tables
writeDatabaseSchema <- "patelm9"
drugExposureIngredientTable <- "ingredient_exposure"
vocabularyTable <- "oncoregimenfinder_vocabulary"

```

### Vocabulary Table  
#### Write    
```{r vocabulary,eval=FALSE,echo=TRUE}
OncoRegimenFinder::createVocabTable(conn = conn,
                 writeDatabaseSchema = writeDatabaseSchema,
                 cdmDatabaseSchema = cdmDatabaseSchema,
                 vocabularyTable = vocabularyTable)

```

#### Row Count History

```{r vocabcount, echo=FALSE,eval=TRUE, print=TRUE}
grep(pattern = vocabularyTable,
    pg13::lsTables(conn = conn,
                   schema = writeDatabaseSchema),
    ignore.case = T,
    value = TRUE) %>% 
        rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = writeDatabaseSchema,
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)

```


#### At a Glance 

```{r samplevocabulary, echo=FALSE, eval=TRUE, cache=TRUE}
vocabularyTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = writeDatabaseSchema,
                                               tableName = vocabularyTable,
                                               n = 20,
                                               n_type = "random"))

vocabularyTableData

```


### Ingredient Exposure Table  

#### Write Ingredient Exposure Table      

```{r ingredientexposures, echo=TRUE,eval=FALSE}

OncoRegimenFinder::buildIngredientExposuresTable(conn = conn,
                                                 cdmDatabaseSchema = cdmDatabaseSchema,
                                                 writeDatabaseSchema = writeDatabaseSchema,
                                                 drugExposureIngredientTable = drugExposureIngredientTable)

```

#### Row Count History   

```{r ingredientexposurescount, echo=FALSE,eval=TRUE}
grep(pattern = drugExposureIngredientTable,
    pg13::lsTables(conn = conn,
                   schema = writeDatabaseSchema),
    ignore.case = T,
    value = TRUE) %>% 
        rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = writeDatabaseSchema,
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)
```

#### At a Glance   

```{r sampleingredientexposure, eval=TRUE, echo=FALSE, cache=TRUE}

drugExposureIngredientTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = writeDatabaseSchema,
                                               tableName = drugExposureIngredientTable,
                                               n = 20,
                                               n_type = "random"))

drugExposureIngredientTableData %>%
  dplyr::select(ingredient_concept_id,
                starts_with("drug_exposure"))

```


```{r dc1, echo=F, eval=TRUE, message="hide"}

fantasia::dcOMOP(conn = conn)

```


# Run Algorithm  
The algorithm can be run once the tables are written in [Setup]. 

## Requirements  
1. A database connection object `conn` to your OMOP CDM instance  
1. `writeDatabaseSchema` destination schema with write privileges  
1. `writeDatabaseSchema`.`vocabularyTable` & `writeDatabaseSchema`.`drugExposureIngredientTable` Tables (see ["Setup"])  

## Outputs  
### Intermediate   
1. Cohort Table `cohortTable`: Ingredient Exposure table filtered for all the RxNorm Ingredients that are descendants of a superclass comprised of one or more OMOP Vocabulary Concept Classes Concept Ids (`drug_classification_id_input`). By default, `drug_classification_id_input` is set to Concept Classes of `OncoRegimenFinder::atc_antineoplastic_id` and `OncoRegimenFinder::hemonc_classes`, but is modifiable according to user preference.  
1. Regimen Staging Table `regimenStagingTable`: A copy of the Cohort Table. This table is written to serve as the data source for running OncoRegimenFinder against a cohort requiring non-standard `date_lag_input` and `regimen_repeats` parameters.   

### Final  
1. *Regimen Table `regimenTable`*: Regimen Staging Table that is run through the OncoRegimenFinder algorithm according to the default `date_lag_input` and `regimen_repeats` values of 30 days and 5, respectively.  
1. *Regimen Ingredient Table `regimenIngredientTable`*: Regimen Ingredient Table that maps the algorithm output in the Regimen Table to HemOnc Regimens in the `vocabularyTable` written in [Setup]. 


## Procedure 
### Load Packages  

```{r pkgsetup2, echo=TRUE,eval=TRUE,message=FALSE,results="hide", warning=FALSE}
library(OncoRegimenFinder)
library(tidyverse)
```

```{r secretpkgsetup2, echo=FALSE,eval=TRUE,message=FALSE,results="hide",warning=FALSE}
library(fantasia)
conn <- fantasia::connectOMOP()
```

### Set Parameters   

```{r setparameters2, eval=TRUE}
# Schema of Source Person and Drug Exposure Tables
cdmDatabaseSchema <- "omop_cdm_2"

# Output Schema and Tables
writeDatabaseSchema <- "patelm9"
drugExposureIngredientTable <- "ingredient_exposure"
cohortTable <- "oncoregimenfinder_cohort"
regimenTable <- "oncoregimenfinder_regimen"
regimenStagingTable <- "oncoregimenfinder_regimen_staging"
vocabularyTable <- "oncoregimenfinder_vocabulary"
regimenIngredientTable <- "oncoregimenfinder_regimen_ingredients"

# OMOP Vocabulary Drug Classes to filter Drug Exposures for
drug_classification_id_input <- c(#ATC Antineoplastics (OncoRegimenFinder::atc_antineoplastics_id)
                                 21601387, 
                                 #HemOnc Classes (OncoRegimenFinder::hemonc_classes)
                                 35101847, 
                                 35807195, 
                                 35807466, 
                                 35807470, 
                                 35807489)
false_positive_id <- 
                #OncoRegimenFinder::falsepositives
                c(45775396, 1304850, 792993, 19010482, 19089602, 
                  19080458, 19104221, 19065450, 1510328, 42903942, 
                  1354698, 1551860, 19003472, 1356009, 40244464, 
                  1303425, 745466, 1389464, 19025194, 740910, 1710612, 
                  35606631, 19003999, 985708, 45775206, 1308432, 1522957, 
                  1760616, 950637, 1300978, 44816310, 1500211, 1506270, 
                  923645, 19061406, 40171288, 1388796, 40168303, 975125, 
                  1507705, 1511449, 1738521, 1548195, 1518254, 40222444, 
                  1713332, 1112807, 19014878, 19034726, 984232, 1525866, 
                  989482, 1550557, 1551099, 924120, 904351)

# Date difference when assessing for drug combinations in the Drug Exposures table
date_lag_input <- 30
regimen_repeats <- 5

```

### Steps  
1. Cohort, Regimen, and a copy of Regimen as "Regimen Staging" Tables are written. 
1. Regimen Table undergoes further processing by combining overlapping instances of an ingredient start date and +/- the date lag input over the number of regimen repeats.  
1. A final Regimen Ingredient Table joins the Vocabulary Table with the Regimen Table to map drug combinations derived from this algorithm back to the HemOnc ontology and other OMOP Concepts  

#### Write Cohort and Regimen Staging Tables

```{r cohortregimentables, eval=FALSE, echo=TRUE}
OncoRegimenFinder::buildCohortRegimenTable(conn = conn,
                                           cdmDatabaseSchema = cdmDatabaseSchema,
                                           writeDatabaseSchema = writeDatabaseSchema,
                                           cohortTable = cohortTable,
                                           regimenTable = regimenTable,
                                           drug_classification_id_input = drug_classification_id_input,
                                           false_positive_id = false_positive_id)
```

##### Cohort Table   
The `Cohort Table` selects for the Person Id, Drug Exposure Id with Start and End Dates for all Drug Concept Ids that are RxNorm Ingredients representing that Drug Exposure filtered for all descendants of the curated Drug Cl argument.  

###### At a Glance   

```{r cohort, eval=TRUE, echo=FALSE, cache=TRUE}

cohortTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = "patelm9",
                                               tableName = "oncoregimenfinder_cohort",
                                               n = 20,
                                               n_type = "random"))

cohortTableData

```

###### Row Count History

```{r cohortcount, echo=FALSE,eval=TRUE}
grep(pattern = "oncoregimenfinder_cohort",
    pg13::lsTables(conn = conn,
                   schema = "patelm9"),
    ignore.case = T,
    value = TRUE) %>% 
        rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = "patelm9",
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)
  

```


##### Regimen Staging Table  
The final Regimen Table is derived from the Regimen Staging Table is also to source of Regimens used to process special cohort tables. The Regimen Staging Table also serves as the data source when running OncoRegimenFinder against special cohorts.  

###### At a Glance  

```{r, echo=FALSE, eval=TRUE,cache=TRUE}

regimenStagingTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = writeDatabaseSchema,
                                               tableName = regimenStagingTable,
                                               n = 20,
                                               n_type = "random"))

regimenStagingTableData

```

###### Row Count History  

```{r regimenstagingcount, echo=FALSE, eval=TRUE}

grep(pattern = regimenStagingTable,
    pg13::lsTables(conn = conn,
                   schema = writeDatabaseSchema),
    ignore.case = T,
    value = TRUE) %>% 
        rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = writeDatabaseSchema,
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)

```


#### Process Regimen Table  
The Regimen Table is grouped by Person Id, Drug Exposure Id, Ingredient and Ingredient Start Date and joined onto itself based on overlapping window of time by +/= the Date Lag Input parameter and iterated on based on the Regimen Repeats given.  

```{r processregimen, eval=FALSE,echo=TRUE}

OncoRegimenFinder::processRegimenTable(conn = conn,
                    writeDatabaseSchema = writeDatabaseSchema,
                    regimenTable = regimenTable,
                    date_lag_input = date_lag_input,
                    regimen_repeats = regimen_repeats)

```

##### Regimen Table  
###### At a Glance  

```{r sampleregimen, echo=FALSE, eval=TRUE, cache=TRUE}

regimenTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = writeDatabaseSchema,
                                               tableName = regimenTable,
                                               n = 20,
                                               n_type = "random"))

regimenTableData

```

###### Row Count History    

```{r regimencount, echo=FALSE, eval=TRUE}
grep(pattern = paste(regimenStagingTable, regimenIngredientTable, sep = "|"),
    pg13::lsTables(conn = conn,
                   schema = writeDatabaseSchema),
    ignore.case = T,
    value = TRUE,
    invert = TRUE) %>% 
  purrr::map(function(x) grep(regimenTable, x, ignore.case = T, value = TRUE)) %>% 
  purrr::keep(~length(.)==1) %>% 
  unlist() %>% 
  rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = writeDatabaseSchema,
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)

```

#### Write Regimen Ingredient Table   

```{r regingr, eval=FALSE, echo = TRUE}
createRegimenIngrTable(conn = conn,
                       writeDatabaseSchema = "patelm9",
                       cohortTable = "oncoregimenfinder_cohort",
                       regimenTable = "oncoregimenfinder_regimen",
                       regimenIngredientTable = "oncoregimenfinder_regimen_ingredients",
                       vocabularyTable = "oncoregimenfinder_vocabulary")
```

##### Regimen Ingredient Table  

###### At a Glance  

```{r sampleregingr, echo=FALSE, eval=TRUE, cache=TRUE}

regimenIngrTableData <-
  pg13::query(conn = conn,
              sql_statement = pg13::buildQuery(schema = writeDatabaseSchema,
                                               tableName = regimenIngredientTable,
                                               n = 20,
                                               n_type = "random"))

regimenIngrTableData

```

###### Row Count History  

```{r regingrcount, echo=FALSE, eval=TRUE}

grep(pattern = regimenIngredientTable,
    pg13::lsTables(conn = conn,
                   schema = writeDatabaseSchema),
    ignore.case = T,
    value = TRUE) %>% 
        rubix::map_names_set(function(x) pg13::query(conn = conn,
                                          pg13::renderRowCount(schema = writeDatabaseSchema,
                                                   tableName = x))) %>% 
        dplyr::bind_rows(.id = "Table") %>% 
        dplyr::rename(RowCount = count)

```


```{r dc2,echo=FALSE,eval=TRUE,message=FALSE,results="hide"}
fantasia::dcOMOP(conn = conn,
                 remove = TRUE)
```

